---
title: "Data Exploration Project: Data Cleaning"
author: "Paul Gano"
date: '2022-04-24'
output: html_document
---
# Loading Libraries

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
library(dplyr)
library(lubridate)
```

# Loading in Datasets

```{r Loading Trends ScoreCard and Ids}
Google_Trends <- list.files(getwd(),pattern = "trends",full.names = TRUE)
Google_Trends <- Google_Trends %>% map_df(read_csv)
Scorecard <- read_csv('Most+Recent+Cohorts+(Scorecard+Elements).csv')
Link_Names <- read_csv('id_name_link.csv')
```
# Reducing to One ID per School

```{r Reducing to One ID per School}
TLink_Names <- Link_Names %>% 
  group_by(schname) %>% 
  mutate(N = n()) %>% 
  filter(N == 1) %>% 
  select(-N)
```

# Joining Data, Filtering Predominantly Bachelor Degrees

```{r Joining data}
Google_Trends_Linked <- Google_Trends %>% inner_join(Link_Names, by = 'schname')
Scorecard_Trends <- Google_Trends_Linked %>% 
  rename(OPEID = opeid) %>% 
  inner_join(Scorecard, by = 'OPEID')
```

```{r filtering to primarily bachelor}
Scorecard_Trends <- Scorecard_Trends %>% filter(PREDDEG == 3)
```

# Adding Months and Weeks, Defining post treatment

```{r Months and Weeks}
string <- str_sub(Scorecard_Trends$monthorweek,1,10)

Scorecard_Trends <- Scorecard_Trends %>% 
  mutate(Week = ymd(string),Month = floor_date(Week, unit = 'month'), After_Score = as.factor(Week > ymd('2015-09-01')))
```

# Standardizing Index

```{r Standardizing Index}
Scorecard_Trends <- Scorecard_Trends %>% 
  group_by(schname, keyword) %>% 
  mutate(Stnd_index = scale(index)) %>% 
  drop_na()
Scorecard_Trends %>% group_by(schname, keyword, Month) %>% summarize(Stnd_index)
```

# Defining High-Earnings

```{r Defining High-Earnings}
Scorecard_Trends$`md_earn_wne_p10-REPORTED-EARNINGS` <- as.numeric(Scorecard_Trends$`md_earn_wne_p10-REPORTED-EARNINGS`)
Scorecard_Trends <- Scorecard_Trends %>% 
  drop_na(`md_earn_wne_p10-REPORTED-EARNINGS`) %>% 
  ungroup() %>% 
  mutate(High_Earning = as.factor(`md_earn_wne_p10-REPORTED-EARNINGS` > mean(`md_earn_wne_p10-REPORTED-EARNINGS`) + mean(`md_earn_wne_p10-REPORTED-EARNINGS`)*.5))
```
I chose to define high earning as 50% more than the mean earnings, I did not want to define high earning as just exclusively above average. That would mean even a school with median earnings $10 above the average would be defined as "high-earning". The choice of 50% makes the breaking point greater than 65k which I think is a reasonable definition of high earning.

# Choosing Final Columns

```{r Choosing Final Columns}
Scorecard_Trends_Final <- Scorecard_Trends %>% 
  select(c(schname, keyword, LOCALE, `md_earn_wne_p10-REPORTED-EARNINGS`, Week, Month, After_Score, Stnd_index, High_Earning))
vtable::vtable(Scorecard_Trends_Final)
```

# Saving Down Final Data

```{r Saving Down Final Data}
write.csv(Scorecard_Trends_Final, "Scorecard_Trends_Final.csv", row.name = TRUE)
```


